%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%                                                                 %
%                         ACKNOWLEDGEMENT                         %
%                                                                 %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
 
\specialhead{ACKNOWLEDGMENT}
 It has been a great privilege to spend two and a half years in the Department of Computer Science at Rensselaer Polytechnic Institute. One of the joys of a completion is to look over the journey past and remember all the people who have helped and supported me along this long but fulfilling road.
 
 My first debt of gratitude must go to my advisor, Dr. Sibel Adal\i. She patiently provided the vision, encouragement and advise necessary for me to proceed the research journey. I hope that I could be as lively, enthusiastic, and energetic as 
Sibel and to someday be able to command an audience as well as she can. Not only that she has always been an inspirational and experienced advisor to me throughout my graduate school years, but also she supports my choice in exploring various subjects with great freedom. I could not have asked for a better role model as a researcher, as well as a great person.


I would like to thank my committee, Dr. Elliot Anshelevich and Dr. Petros Drineas for their support and guidance. Special thanks to Dr. Elliot Anshelevich to his many valuable suggestions. The guidance has served me well and I owe them my heartfelt appreciation.
 
I wish to thank my parents, Guahua Qian and Pinhong Jiang. Their love provided my inspiration and was my driving force through the years. My girl, Feifei Bian, whose love and encouragement brings me all the way here. In our four-year relationship, we bear the brunt of frustration and share the joy of success.  Without the love and support of them, I would certainly not be able to finish this thesis.

I also would like to thank T. DuBois for sharing the implementation details of his algorithm.
