%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%                                                                 %
%                           BIBLIOGRAPHY                          %
%                                                                 %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
 
%This method produces a numbered bibliography where the numbers
%correspond to the \cite commands in the text. See the LaTeX manual.
%
\specialhead{LITERATURE CITED}
\begin{singlespace}
\begin{thebibliography}{99}
\bibitem{thisbook} This is the first item in the Bibliography.
Let's make it very long so it takes more than one line.
Let's make it very long so it takes more than one line.
Let's make it very long so it takes more than one line.
Let's make it very long so it takes more than one line. 
\bibitem{anotherbook} The second item in the Bibliography.
\bibitem{yetanotherbook} Another item in the Bibliography.
\end{thebibliography}
\end{singlespace}
   
%This is an alternative method.  It's a simple unnumbered bibliography
%with hanging indentation.
%
\specialhead{BIBLIOGRAPHY}
\begin{singlespace}
\bibentry This is the first item in the Bibliography.
Let's make it very long so it takes more than one line.
Let's make it very long so it takes more than one line.
Let's make it very long so it takes more than one line.
Let's make it very long so it takes more than one line.
Let's make it very long so it takes more than one line. 
\bibentry The second item in the Bibliography.
\bibentry Another item in the Bibliography.
  
\end{singlespace} 

% This file was created with JabRef 2.5.
% Encoding: UTF-8

@incollection{Ames:2011,
author = "D.L. Ames and S.T. Fiske and A.T Todorov",
title = "Impression Formation: A Focus on Others’ Intents",
year = "2011",
booktitle = "The Oxford Handbook of Social Neuroscience",
editor = "J. Decety and J. Cacioppo",
publisher = "Oxford University Press",
pages = "419-433"
}

@ARTICLE{Fiske:2007,
title = "Universal dimensions of social cognition: warmth and competence",
journal = "Trends in Cognitive Sciences",
volume = "11",
number = "2",
pages = "77 - 83",
year = "2007",
note = "",
issn = "1364-6613",
doi = "10.1016/j.tics.2006.11.005",
url = "http://www.sciencedirect.com/science/article/pii/S1364661306003299",
author = "Susan T. Fiske and Amy J.C. Cuddy and Peter Glick"
}

@BOOK{Adali:2013,
  title = {Modeling Trust Context in Networks},
  publisher = {Springer Briefs},
  year = {2013},
  author = {Sibel Adal{\i}}
}

@ARTICLE{Cartwright:56,
  author = {D. Cartwright and F. Harary},
  title = {Structural balance: a generalization of Heider's theory},
  journal = {Psychological Review},
  year = {1956},
  volume = {63},
  pages = {277-293},
  number = {5}
}

@ARTICLE{Davis:67,
  author = {J. Davis},
  title = {Clustering and structural balance in graphs},
  journal = {Human Relations},
  year = {1967},
  pages = {181-187}
}

@ARTICLE{Doreian:02,
  author = {P. Doreian},
  title = {Event sequences as generators of social network evolution},
  journal = {Social Networks},
  year = {2002},
  volume = {24},
  pages = {93-119},
  owner = {qiany3},
  timestamp = {2013.03.01}
}

@INPROCEEDINGS{golbeck:distrust2011,
  author = {T. DuBois and J. Golbeck and A. Srinavasan},
  title = {Predicting Trust and Distrust in Social Networks},
  booktitle = {IEEE International Conference on Social Computing},
  year = {2011}
}

@INPROCEEDINGS{DuBois:2009,
  author = {T. DuBois and J. Golbeck and A. Srinivasan},
  title = {Rigorous probabilistic trust-inference with applications to clustering},
  booktitle = {Proceedings of the 2009 IEEE/WIC/ACM International Joint Conference
	on Web Intelligence and Intelligent Agent Technology},
  year = {2009},
  volume = {01},
  pages = {655–658},
  publisher = {IEEE Computer Society}
}

@BOOK{kleinberg-book,
  title = {Networks, Crowds, and Markets: Reasoning About a Highly Connected
	World},
  publisher = {Cambridge University Press},
  year = {2010},
  author = {David Easley and Jon Kleinberg}
}

@ARTICLE{Gansner:05,
  author = {E. R. Gansner and Y. Koren and S. North},
  title = {Graph Drawing by Stress Majorization},
  journal = {J. Pach (Ed.), Graph Drawing, Lecture Notes in Computer Science},
  year = {2005},
  volume = {3383},
  pages = {239-250},
  publisher = {Berlin/Heidelberg: Springer}
}

@ARTICLE{Granovetter:1973,
  author = {M. Granovetter},
  title = {The Strength of Weak Ties},
  journal = {American Journal of Sociology},
  year = {1973},
  volume = {78},
  pages = {1-22},
  issue = {6}
}

@INPROCEEDINGS{Guha:04,
  author = {R. Guha and R. Kumar and P. Raghavan and A. Tomkins},
  title = {Propagation of trust and distrust},
  booktitle = {Proceedings of the 13th international conference on World Wide Web.
	ACM},
  year = {2004},
  pages = {403--412}
}

@ARTICLE{Heider:46,
  author = {F. Heider},
  title = {Attitudes and cognitive organization},
  journal = {Journal of Psychology},
  year = {1946},
  pages = {107-12}
}

@INPROCEEDINGS{Khoury:12,
  author = {M. Khoury and Yifan Hu and Shankar Krishnan and Carlos Scheidegger},
  title = {Drawing Large Graphs by Low-Rank Stress Majorization},
  booktitle = {EuroVis},
  year = {2012}
}

@INPROCEEDINGS{Leskovec:2010,
  author = {J. Leskovec and D. Huttenlocher and J. Kleinberg},
  title = {Predicting positive and negative links in online social networks},
  booktitle = {Proceedings of the 19th International Conference on World Wide Web},
  year = {2010},
  pages = {641--650}
}

@INPROCEEDINGS{Kleinberg:03,
  author = {D. Liben-Nowell and J. Kleinberg},
  title = {The link prediction problem for social networks},
  booktitle = {International Conference on Information and Knowledge Management
	(CIKM)},
  year = {2003},
  pages = {556-559}
}

@ARTICLE{Tomasello:2005,
  author = {Tomasello,Michael and Carpenter,Malinda and Call,Josep and Behne,Tanya
	and Moll,Henrike},
  title = {Understanding and sharing intentions: The origins of cultural cognition},
  journal = {Behavioral and Brain Sciences},
  year = {2005},
  volume = {28},
  pages = {675-91; discussion 691-735},
  number = {5}
}

@ARTICLE{Uzzi:1996,
  author = {Brian Uzzi},
  title = {The sources and consequences of embeddedness for the economic performance
	of organizations: The network effect.},
  journal = {American sociological review},
  year = {1996},
  volume = {61},
  pages = {674-698},
  issue = {4}
}




% Note that, if you wish, you can use BibTeX to create your bibliography
% from a database. See section 5.6.2 of Memo RPI.110 for information. 
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: 
